Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-02-19 resex.tex
Started revision of HEL section and of bibliography. Renamed bibliography (now managed with BibDesk).

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage{color}
\usepackage{siunitx}
\usepackage{url}

@@

In a recent review, the need for such an active halo control system
for HL-LHC has been assessed, with the conclusion that it would
considerably increase the operational margins and reduce the risks for
machine protection~\cite{helreview}. In view of the need of active
halo control for HL-LHC and for future high power accelerators, like
HE-LHC and FCC-hh~\cite{helhcparam2011, fcc_coll_ipac2017}, different
active halo control methods have been
studied~\cite{helreview_bruce}. The hollow electron lens (HEL) is
considered the most established, flexible and suitable technology for
the HL-LHC~\cite{helreview}.

@@

\subsection{General Overview\label{sec:hel:intro}}
Electron lenses produce DC or pulsed low energy electron beams, where the electron beam is generated in an electron gun, then guided and confined along a straight section by strong solenoids and finally dumped onto a collector. As an example the conceptual design of the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
\begin{figure}[h]
	\includegraphics[width=1.0\linewidth]{hel_layout_epbeam}% Here is how to import EPS art
	\caption{\label{fig:hel_layout} Layout of HL-LHC HEL. Courtesy of CERN EN-MME group.}
\end{figure}

@@

\subsection{Operation modes and effects on the beam core\label{sec:hel:core}}
For the HEL two modes of operation are currently under consideration, the DC mode as standard operation mode and the pulsed mode. The main benefit of pulsed HEL operation is the increase in halo removal rates which might become essential as in ideal operation conditions featuring low non-linearities, in particular low chromaticity and octupole current, the DC mode is not effective~\cite{hel_halo_hllhc_fitterer,hl_halo_ipac2017}. 
For this purpose, two different pulsing patterns are thus considered for the HL-LHC:
\begin{itemize}
	\item \textbf{random excitation:} The electron gun voltage is modulated between:
	\begin{eqnarray}
	U_{\mathrm{e-gun}}&=&a\cdot U_{\mathrm{max}}\\
	& &+(1-a)\cdot \mathrm{ran}(0,1)\cdot U_{\mathrm{max}}
	\end{eqnarray}
	with $U_{\mathrm{max}}$ the maximum voltage in DC operation, $a$ the modulation strength with $a\in[0,1]$, and $\mathrm{ran}(0,1)$ a uniformly distributed random number between~[0,1].
	\item \textbf{resonant excitation:} The HEL is switched on only every $k^{\mathrm{th}}$ turn. The excitation can then be represented by:
	\begin{eqnarray}\label{intro:eqn:1}
	f(t)&=&\sum_{n=-\infty}^{+\infty}\delta(t-n\cdot(kT)),
	\end{eqnarray}
	where $n$ is the turn number and $T$ the revolution time, and its Fourier series by:
	\begin{eqnarray}\label{intro:eqn:2}
	f(t)=\Sh_{kT}(t)
	&=&\frac{1}{kT}\sum_{n=-\infty}^{+\infty}e^{2\pi i f_nt} \\
	& & \quad \text{with} \ f_n=\frac{n}{k}f_{\rm rev},
	\end{eqnarray}
	where $\Sh_{kT}$ is the Dirac comb. As can be seen from the Fourier series, $k^{\mathrm{th}}$~turn pulsing in general drives $k^{\mathrm{th}}$ order resonances \cite{md_sim_hel_res_ex_fitterer}. Historically, the $k^{\mathrm{th}}$~turn pulsing was used in regular operation in the Tevatron for abort cleaning \cite{hel_tevatron_abortgap_zhang}.
\end{itemize}

@@

%\section{Additional plots for random excitation during 2017 experiments}
%\label{app:sec:ran}

@@

\bibliography{bibliography}% Produces the bibliography via BibTeX.

\end{document}
%
% ****** End of file apssamp.tex ******

>>>>

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage{color}
\usepackage{siunitx}
\usepackage{url}

@@

In a recent review, the need for such an active halo control system
for HL-LHC has been assessed, with the conclusion that it would
considerably increase the operational margins and reduce the risks for
machine protection~\cite{helreview}. In view of the need of active
halo control for HL-LHC and for future high power accelerators, like
HE-LHC and FCC-hh~\cite{helhcparam2011, fcc_coll_ipac2017}, different
active halo control methods have been
studied~\cite{helreview_bruce}. The hollow electron lens (HEL) is
considered the most established, flexible and suitable technology for
the HL-LHC~\cite{hel_tevatron_stancari, helreview}.

@@

\subsection{General Overview}
\label{sec:hel:intro}

Electron lenses are based upon continuous or pulsed low energy,
magnetically confined electron
beams~\cite{Shiltsev:elens-book:2016}. The electron beam is generated
in an electron gun, guided and confined by strong solenoids and
finally dumped in a collector. As an example, the conceptual design of
the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.

\begin{figure*}
  \includegraphics[width=0.75\textwidth]{hel_layout_epbeam}
  \caption{Layout of the hollow electron lens for HL-LHC. (Courtesy of
    CERN EN-MME mechanical engineering group.)}
  \label{fig:hel_layout}
\end{figure*}

@@

\subsection{Operation modes and effects on the beam core}
\label{sec:hel:core}

For the HEL two modes of operation are currently under consideration,
the continuous (also referred to as `DC', or direct current, in this
paper) mode as standard operation mode and the pulsed mode. The main
benefit of pulsed HEL operation is the increase in halo removal rates
which might become essential as in ideal operation conditions
featuring low non-linearities, in particular low chromaticity and
octupole current, the DC mode is not
effective~\cite{hel_halo_hllhc_fitterer,hl_halo_ipac2017}.

For this purpose, two different pulsing patterns are thus considered
for the HL-LHC:

\begin{itemize}
	\item \textbf{random excitation:} The electron gun voltage is modulated between:
	\begin{eqnarray}
	U_{\mathrm{e-gun}}&=&a\cdot U_{\mathrm{max}}\\
	& &+(1-a)\cdot \mathrm{ran}(0,1)\cdot U_{\mathrm{max}}
	\end{eqnarray}
	with $U_{\mathrm{max}}$ the maximum voltage in DC operation, $a$ the modulation strength with $a\in[0,1]$, and $\mathrm{ran}(0,1)$ a uniformly distributed random number between~[0,1].
	\item \textbf{resonant excitation:} The HEL is switched on only every $k^{\mathrm{th}}$ turn. The excitation can then be represented by:
	\begin{eqnarray}\label{intro:eqn:1}
	f(t)&=&\sum_{n=-\infty}^{+\infty}\delta(t-n\cdot(kT)),
	\end{eqnarray}
	where $n$ is the turn number and $T$ the revolution time, and its Fourier series by:
	\begin{eqnarray}\label{intro:eqn:2}
	f(t)=\Sh_{kT}(t)
	&=&\frac{1}{kT}\sum_{n=-\infty}^{+\infty}e^{2\pi i f_nt} \\
	& & \quad \text{with} \ f_n=\frac{n}{k}f_{\rm rev},
	\end{eqnarray}
	where $\Sh_{kT}$ is the Dirac comb. As can be seen from the Fourier series, $k^{\mathrm{th}}$~turn pulsing in general drives $k^{\mathrm{th}}$ order resonances \cite{md_sim_hel_res_ex_fitterer}. Historically, the $k^{\mathrm{th}}$~turn pulsing was used in regular operation in the Tevatron for abort cleaning \cite{hel_tevatron_abortgap_zhang}.
\end{itemize}

@@

@@

\bibliography{resex-bibdesk}

\end{document}
